\name{}: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-04-11 paper/main.tex
.

\begin{figure}[!tpb]%figure1
	\centerline{\includegraphics[width=1\columnwidth]{figures/full_model_all_layers.png}}
	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
	\label{fig:FullModelAllLayers}
\end{figure}

\subsubsection{Combined prediction model}
Combining half-twin and graph convolutional neural networks, we map
all protein representations to their respective node features,
initializing the graph convolutional update steps. The resulting
representations are used for a similarity prediction.
% as presented in
% figure ~\ref{fig:HalfTwinNetwork}
When combining ontology and molecular features with or without the
graph model, we concatenate both protein features and both drugs
features, before plugging them into the graph model for the similarity
computation. An overview of the model architecture, combining both
feature types, is shown in Figure ~\ref{fig:FullModelAllLayers}. Here
the original representations are transformed by a dense layer and then
used as input of a stack (with height 3) of residual graph
convolutional blocks.

@@

To find the best hyperparameter configuration for the proposed model,
we performed a grid search to find the most expressive and
non-redundant representation. We pretrained the bottom-up and the
top-down model separately and aimed at best performing models
with respect to our evaluation metrics. We optimized embedding sizes,
depth of the neural network, optimizer, learning rate and layer types
using an extensive, manual grid search. Starting from na\"ive, shallow
feature transformations with an embedding size of $10$, we scaled the
network up to residual structures with up to $10$ hidden layers
leading to embeddings of size $4000$, testing different network widths
and learning rates for each configuration.

@@

We trained, validated and finally tested all considered models on the
STITCH dataset using a 5-fold cross-validation over a protein split;
we then selected the best-performing models (with respect to
$MicroAUC_p$, see Section \ref{sec:evaluation}) and evaluate them in a
5-fold protein-split cross-validation on the Yamanishi benchmark
dataset to avoid validation overfitting and yield more realistic
testing results. To evaluate the influence of the different features
separately, and to determine whether they ``localize'' on the PPI
graph (and therefore can be exploited successfully by the graph neural
networks), we train and evaluate models with different types of
features, and with and without inclusion of the PPI graph,
separately. We comparing the molecular (MolPred) and phenotype-based
(OntoPred) prediction model, and a combination of both where we
concatenate both types of features.  Table \ref{tab:Results} shows the
results of these experiments.

>>>>

% \begin{figure}[!tpb]%figure1
% 	\centerline{\includegraphics[width=1\columnwidth]{figures/full_model_all_layers.png}}
% 	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
% 	\label{fig:FullModelAllLayers}
% \end{figure}

\subsubsection{Combined prediction model}
Combining half-twin and graph convolutional neural networks, we map
all protein representations to their respective node features,
initializing the graph convolutional update steps. The resulting
representations are used for a similarity prediction.
% as presented in
% figure ~\ref{fig:HalfTwinNetwork}
When combining ontology and molecular features with or without the
graph model, we concatenate both protein features and both drugs
features, before plugging them into the graph model for the similarity
computation. An overview of the model architecture, combining both
feature types, is shown in Supplementary Figure
\ref{fig:FullModelAllLayers}\todo{Move to supplement}. Here
the original representations are transformed by a dense layer and then
used as input of a stack (with height 3) of residual graph
convolutional blocks.

@@

To find the best hyperparameter configuration for the proposed model,
we performed a grid search to find the most expressive and
non-redundant representation. We pretrained the bottom-up and the
top-down model separately and aimed at best performing models with
respect to our evaluation metrics. We optimized embedding sizes, depth
of the neural network, optimizer, learning rate and layer types using
an extensive, manual grid search. Starting from shallow feature
transformations with an embedding size of $10$, we scaled the network
up to residual structures with up to $10$ hidden layers leading to
embeddings of size $4000$, testing different network widths and
learning rates for each configuration.

@@

We trained, validated and finally tested all considered models on the
STITCH dataset using a 5-fold cross-validation over a protein split;
we then selected the best-performing models (with respect to
$MicroAUC_p$, see Section \ref{sec:evaluation}) and evaluate\todo{Only
evaluate or retrain (from scratch)? Should say here.} them in a
5-fold protein-split cross-validation on the Yamanishi benchmark
dataset to avoid validation overfitting and yield more realistic
testing results. To evaluate the influence of the different features
separately, and to determine whether they ``localize'' on the PPI
graph (and therefore can be exploited successfully by the graph neural
networks), we train and evaluate models with different types of
features, and with and without inclusion of the PPI graph,
separately. We comparing the molecular (MolPred) and phenotype-based
(OntoPred) prediction model, and a combination of both where we
concatenate both types of features.  Table \ref{tab:Results} shows the
results of these experiments.
